Wavelet Multiresolution Spatio-Temporal Modelling Using the Integro-Difference Equation

### 2010-11-08 Multiresolution_framework/ltx/manuscript.tex
added bib file
edits to estimation section

%    \documentclass[journal,a4paper]{IEEEtran}
%     \documentclass[10pt,twocolumn,twoside]{IEEEtran}
     \documentclass[11pt,draftcls,onecolumn,peerreview]{IEEEtran} 
%         \topmargin       -6.0mm
%          \oddsidemargin      0mm
%          \evensidemargin     0mm
%          \textheight     223.5mm
%         \textwidth      170.0mm

@@

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals use top floats exclusively.
% Note that, LaTeX2e, unlike IEEE journals, places footnotes above bottom
% floats. This can be corrected via the \fnbelowfloat command of the
% stfloats package.
\section{MRA of the IDE in State-Space}
The spatially homogeneous, linear IDE is given by
\begin{equation}
 z_{t+1}\left(\mathbf{s}\right)=\int_{\mathcal{S}}k\left(\mathbf{s}-\mathbf{r}\right)z_{t}\left(\mathbf{r}\right)d\mathbf{r}+e_{t}\left(\mathbf{s}\right)
\label{eq:ConvolutionIntegral}
\end{equation}
where $t\in \mathbb{Z}_0 $ denotes discrete time, $\mathbf{s},\mathbf{r} \in \mathcal{S} \subset \mathbb{R}^{n_\mathcal{S}}$ are spatial locations in $n_{\mathcal S}$-dimensional physical space, where $n_{\mathcal S} \in \left\lbrace 1,2,3 \right\rbrace $. The continuous spatial field at time $t$ and at location $\mathbf s$ is denoted $z_t\left(\mathbf s\right)$. The model dynamics  are defined by the homogeneous time invariant spatial mixing kernel, $k\left(\mathbf{s}-\mathbf{r}\right):\mathbb{R}^{n_\mathcal{S}}\rightarrow \mathbb{R}$ which maps the current spatial field to the next spatial field via the integral (\ref{eq:ConvolutionIntegral}). The disturbance $e_{t}(\mathbf{s})$ is a zero-mean normally distributed noise process, spatially coloured but temporally independent, with covariance 
\begin{equation}
cov\left(e_{t}\left(\mathbf{s}\right),e_{t+\tau}\left(\mathbf{r}\right)\right)=
\begin{cases}
\eta\left(\mathbf{s}-\mathbf{r}\right), & \tau=0 \\
0 & \mathrm{otherwise}
\end{cases}
\label{eq:FieldCovariance}
\end{equation}
for any $\tau \in \mathbb{Z}$, where $\eta\left(\mathbf s-\mathbf r \right)$ is a  spatially homogeneous covariance function. The spatial field is observed by
\begin{equation}
 \mathbf{y}_{t}=\mathbf{z}_{t}+\mathbf{v}_t
\label{eq:observation equation}
\end{equation}
where the observation vector $\mathbf{y}_{t}=[y_t(\mathbf{s}_1) y_t(\mathbf{s}_2)\cdots y_t(\mathbf{s}_{n_y})]^\top$, compiled at $n_{y}$ spatial locations and at time $t$, is corrupted by an i.i.d normally distributed zero-mean white noise $\mathbf{v}_{t}\sim \mathcal{N}\left(\mathbf{0},\mathbf{\Sigma}_{v}\right)$ with $\mathbf{\Sigma}_{v}=\sigma_v^2\mathbf I_{n_y} $. It is assumed here that $\boldsymbol\Sigma_v$ and $\eta\left(\mathbf s-\mathbf r \right)$ are known. The superscript $\top$ denotes the transpose operator.

@@

The state-space representation of the spatio-temporal IDE proposed in \cite{Dewar2009} is derived for an arbitrary set of basis functions and hence this derivation can be applied using scaling and wavelet basis functions. Let
\begin{equation}
 \mathbf{\Lambda}_{x}=\int_{\mathcal{S}}\boldsymbol\mu\left(\mathbf{s}\right)\boldsymbol\mu\left(\mathbf{s}\right)^\top d\mathbf{s}
\label{eq:ConstructionIntegral1}
\end{equation}
\begin{equation}
 \mathbf{\Lambda}_{\theta}=\iint\limits_{\mathcal{S}}\boldsymbol\mu\left(\mathbf{s}\right)\boldsymbol\theta^\top \boldsymbol\lambda\left(\mathbf{s-r}\right)\boldsymbol\mu\left(\mathbf{r}\right)^\top d\mathbf{r}d\mathbf{s}
\label{eq:ConstructionIntegral2}
\end{equation}
where $\boldsymbol\Lambda_{x}, \boldsymbol\Lambda_{\theta} \in \mathbb{R}^{n_x \times n_x}$. The spatial IDE model in (\ref{eq:ConvolutionIntegral}) and (\ref{eq:observation equation}) can be represented as linear state-space model of the form  
\begin{equation}
 \mathbf{x}_{t+1}=\mathbf{A}\left(\boldsymbol\theta\right)\mathbf{x}_t+\mathbf{w}_t
\end{equation}
\begin{equation}
\mathbf{y}_t=\mathbf{C}\mathbf{x}_t+\mathbf{v}_t
\end{equation}
where 
\begin{equation}
 \mathbf{A}\left(\boldsymbol\theta\right)=\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda_{\theta}}
\label{eq:TransitionMatrix}
\end{equation}
and  \\ \\ 
$\mathbf{C}=$
$\begin{pmatrix}
\boldsymbol\phi_{j_0}\left(\mathbf{s}_1\right) & \boldsymbol\phi_{j_0}\left(\mathbf{s}_2\right)  &\cdots &\boldsymbol\phi_{j_0}\left(\mathbf{s}_{n_y}\right) \\
\boldsymbol\psi_{j_0}\left(\mathbf{s}_1\right) & \boldsymbol\psi_{j_0}\left(\mathbf{s}_2\right)  &\cdots &\boldsymbol\psi_{j_0}\left(\mathbf{s}_{n_y}\right) \\
\boldsymbol\psi_{j_0+1}\left(\mathbf{s}_1\right) & \boldsymbol\psi_{j_0+1}\left(\mathbf{s}_2\right)  &\cdots &\boldsymbol\psi_{j_0+1}\left(\mathbf{s}_{n_y}\right) \\
\vdots & \vdots & \ddots& \vdots \\
\boldsymbol\psi_{j}\left(\mathbf{s}_1\right) & \boldsymbol\psi_{j}\left(\mathbf{s}_2\right)  &\cdots &\boldsymbol\psi_{j}\left(\mathbf{s}_{n_y}\right) \\
\end{pmatrix}^{\top}$ \\ \\
\\ \\ and where 
\begin{equation}
 \mathbf{w}_t=\mathbf{\Lambda}_{x}^{-1}\int_{\mathcal{S}}\boldsymbol\mu \left(\mathbf{s}\right)e_t\left(\mathbf{s}\right)d\mathbf{s}
\label{CovofField}
\end{equation}
 is a vector valued, zero-mean normally distributed white noise process with covariance given by (Lemma \ref{lemma:FieldCovariance}, Appendix)
\begin{equation}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\boldsymbol\Pi\mathbf{\Lambda}_{x}^{-\top}
\end{equation}
and
\begin{equation}
 \boldsymbol\Pi=\iint\limits_{\mathcal{S}}\boldsymbol\mu\left(\mathbf{s}\right)\eta\left(\mathbf{s}-\mathbf{r}\right)\boldsymbol\mu\left(\mathbf{r}\right)^{\top}d\mathbf{r}d\mathbf{s}
\label{eq:Pi}
\end{equation}
By substituting the  parameter vector $\boldsymbol\theta$, the  state vector $\mathbf{x_t}$, the vector of kernel basis functions $\boldsymbol\lambda$ and the vector of field basis functions $\boldsymbol\mu$ in (\ref{eq:ConstructionIntegral1}-\ref{eq:Pi}), the state-space representation of the multiresolution approximation of the IDE can be achieved. It should be noted that the complexity of the model is independent of the number of observation locations, providing more flexibility in handling high resolution data. %The model is also capable of being applicable to scenarios with irregular observation locations.

@@

By exploiting properties (\ref{eq:BsplineConvolution}) and (\ref{eq:BsplineInnerProduct}) the  integrals (\ref{eq:ConstructionIntegral1}) and (\ref{eq:ConstructionIntegral2}) can be computed. In order to calculate elements of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Lambda_{\theta}$, scaling and wavelet basis functions should be expanded  in terms of $N_m$ at the appropriate scale. It is important to note that B-spline scaling and wavelet functions possess the following orthogonality properties \cite{Unser1993}: 
\begin{equation}
 \left\langle \psi_{m;j_1,l_1}(s),\psi_{m;j_2,l_2}(s)\right\rangle =0  \quad \mathrm{for} \quad j_1\neq j_2
\label{PsiPsiOrthogonality}
\end{equation}
\begin{equation}
 \left\langle \phi_{m;j_1,l_1}(s),\psi_{m;j_2,l_2}(s)\right\rangle =0  \quad \mathrm{for} \quad j_1\leq j_2
\label{PhiPsiOrthogonality}
\end{equation}
Where the subscript $m$ is introduced to $\phi_{j,l}(s)$ and $\psi_{j,l}(s) $ to indicate the $m$th order B-spline scaling and wavelet functions. In the two dimensional case, the convolution integral can be factorised as follows, e.g. 
\setlength{\arraycolsep}{0.14em}
\begin{IEEEeqnarray}{!l l }
 \phi_{m;j,\mathbf{l}}\left(s_{1},s_{2}\right)\ast \psi^{1}_{m;j,\mathbf{l'}}\left(s_{1},s_{2}\right)&=\left[\phi_{m;j,l_1}\left(s_{1}\right)\ast\phi_{m;j,l'_1}\left(s_{1}\right)\right] \nonumber \\
&\times \left[ \phi_{m;j,l_2}\left(s_{2}\right) \ast \psi_{m;j,l'_2}\left(s_{2}\right)\right] \nonumber \\
&
\end{IEEEeqnarray}
\setlength{\arraycolsep}{5pt}
The preceding factorisation can be also applied to the 2D-inner product integral. Non-zero quantities can  be then expressed by
\setlength{\arraycolsep}{0.14em}
\begin{IEEEeqnarray}{!l l }
\left\langle \phi_{m;j,\mathbf{l}}\left(s_{1},s_{2}\right),\phi_{m;j,\mathbf{l'}}\left(s_{1},s_{2}\right)\right\rangle &=\left\langle \phi_{m;j,l_1}\left(s_{1}\right),\phi_{m;j,l'_1}\left(s_{1}\right)\right\rangle \nonumber \\
& \times\left\langle \phi_{m;j,l_2}\left(s_{2}\right) , \phi_{m;j,l'_2}\left(s_{2}\right)\right\rangle \nonumber \\
&
\end{IEEEeqnarray}
\setlength{\arraycolsep}{5pt}
likewise we can compute 
$\left\langle  \psi^{1}_{m;j,\mathbf{l}}\left(s_{1},s_{2}\right),\psi^{1}_{m;j,\mathbf{l'}}\left(s_{1},s_{2}\right)\right\rangle$
, \\
$\left\langle  \psi^{2}_{m;j,\mathbf{l}}\left(s_{1},s_{2}\right),\psi^{2}_{m;j,\mathbf{l'}}\left(s_{1},s_{2}\right)\right\rangle$ and \\
$
\left\langle \psi^{3}_{m;j,\mathbf{l}}\left(s_{1},s_{2}\right), \psi^{3}_{m;j,\mathbf{l'}}\left(s_{1},s_{2}\right)\right\rangle $. \\ \\
Using these expansions it becomes straightforward to calculate the integrals (\ref{eq:ConstructionIntegral1}) and (\ref{eq:ConstructionIntegral2}). In this work, 4th order cardinal B-spline scaling and wavelet functions are used therefore, to compute the integrals (\ref{eq:ConstructionIntegral1}) and (\ref{eq:ConstructionIntegral2})  the $8$th and $12$th order B-spline values at integer points are  required and can be found in \cite{Goswami1999}.
\section{Estimation of the Multiresolution Spatio-temporal IDE Model}
A state-space representation of the IDE model allows the use of the Expectation Maximisation (EM) algorithm to infer both the kernel and the spatial field from the noise corrupted spatio-temporal data. In this context, the EM algorithm is a two part iterative algorithm that yields the maximum likelihood estimate of the spatial mixing kernel parameters and the posterior distribution of the field weights. In the M-step a lower bound $\mathcal{Q}\left(\boldsymbol\theta,\boldsymbol\theta'\right)$ of the complete-data log-likelihood  $p\left(\mathbf{X},\mathbf{Y} \mid \boldsymbol\theta \right)$, is maximised under the assumption of the system parameters being $\boldsymbol\theta'$ \cite{Gibsona2005}. This can be achieved for the state-space representation of the IDE model via Theorem 2 of \cite{Dewar2009} under very mild conditions.

@@

This section demonstrates the identification of the proposed IDE model. Primarily, the MRA of a synthetic model is demonstrated, which is able to represent the system at different spatial scales. In addition, the complexity of the field decomposition is shown to be independent to the number and spatial regularity of the observation locations. This allows the construction of parsimonious models using spatially dense observations, and demonstrates the utility of the approach to situations where the observations are not arranged on a regular lattice, i.e. sensor networks \cite{Sahu2008}.

@@

@@

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
  \newpage
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{IEEEabrv,../../../Papers/SpatioTemporalSystems,../../../Papers/Wavelet}

>>>>

 % TO DELETE
\newcommand{\mike}[1]{\textsf{\emph{\textbf{\textcolor{red}{#1}}}}} 
\newcommand{\dean}[1]{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}} 
\newcommand{\parham}[1]{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}} 
\newcommand{\ken}[1]{\textsf{\emph{\textbf{\textcolor{magenta}{#1}}}}} 
%    \documentclass[journal,a4paper]{IEEEtran}
%     \documentclass[10pt,twocolumn,twoside]{IEEEtran}
     \documentclass[11pt,draftcls,onecolumn,peerreview]{IEEEtran} 
%         \topmargin       -6.0mm
%          \oddsidemargin      0mm
%          \evensidemargin     0mm
%          \textheight     223.5mm
%         \textwidth      170.0mm

@@

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals use top floats exclusively.
% Note that, LaTeX2e, unlike IEEE journals, places footnotes above bottom
% floats. This can be corrected via the \fnbelowfloat command of the
% stfloats package.
\section{MRA of the IDE in State-Space}
The spatially homogeneous, linear IDE is given by
\begin{equation}
 z_{t+1}\left(\mathbf{s}\right)=\int_{\mathcal{S}}k\left(\mathbf{s}-\mathbf{r}\right)z_{t}\left(\mathbf{r}\right)d\mathbf{r}+e_{t}\left(\mathbf{s}\right)
\label{eq:ConvolutionIntegral}
\end{equation}
where $t\in \mathbb{Z}_0 $ denotes discrete time, $\mathbf{s},\mathbf{r} \in \mathcal{S} \subset \mathbb{R}^{n_\mathcal{S}}$ are spatial locations in $n_{\mathcal S}$-dimensional physical space, where $n_{\mathcal S} \in \left\lbrace 1,2,3 \right\rbrace $. The continuous spatial field at time $t$ and at location $\mathbf s$ is denoted $z_t\left(\mathbf s\right)$. The model dynamics  are defined by the homogeneous time invariant spatial mixing kernel, $k\left(\mathbf{s}-\mathbf{r}\right):\mathbb{R}^{n_\mathcal{S}}\rightarrow \mathbb{R}$ which maps the current spatial field to the next spatial field via the integral (\ref{eq:ConvolutionIntegral}) and $e_t(\mathbf s)$ is an $i.i.d.$ disturbance such that $e_t(\mathbf{s})\sim\mathcal{GP}(\mathbf 0,\eta(\mathbf{s}-\mathbf{r}))$.  Here $\mathcal{GP}(\mathbf 0,\eta(\mathbf{s}-\mathbf{r}))$  denotes a zero mean spatial Gaussian process with covariance function $\eta(\mathbf{s}-\mathbf{r})$~\cite{Rasmussen2005}. The spatial field is observed by
\begin{equation}
 \mathbf{y}_{t}=\mathbf{z}_{t}+\mathbf{v}_t
\label{eq:observation equation}
\end{equation}
where the observation vector $\mathbf{y}_{t}=[y_t(\mathbf{s}_1) y_t(\mathbf{s}_2)\cdots y_t(\mathbf{s}_{n_y})]^\top$, compiled at $n_{y}$ spatial locations and at time $t$, is corrupted by an i.i.d normally distributed zero-mean white noise $\mathbf{v}_{t}\sim \mathcal{N}\left(\mathbf{0},\mathbf{\Sigma}_{v}\right)$ with $\mathbf{\Sigma}_{v}=\sigma_v^2\mathbf I_{n_y} $. It is assumed here that $\boldsymbol\Sigma_v$ and $\eta\left(\mathbf s-\mathbf r \right)$ are known. The superscript $\top$ denotes the transpose operator.

@@

The state-space representation of the spatio-temporal IDE proposed in \cite{Dewar2009} is derived for an arbitrary set of basis functions and hence this derivation can be applied using scaling and wavelet basis functions. Let
\begin{equation}
 \mathbf{\Lambda}_{x}=\int_{\mathcal{S}}\boldsymbol\mu\left(\mathbf{s}\right)\boldsymbol\mu\left(\mathbf{s}\right)^\top d\mathbf{s}
\label{eq:ConstructionIntegral1}
\end{equation}
\begin{equation}
 \mathbf{\Lambda}_{\theta}=\iint\limits_{\mathcal{S}}\boldsymbol\mu\left(\mathbf{s}\right)\boldsymbol\theta^\top \boldsymbol\lambda\left(\mathbf{s-r}\right)\boldsymbol\mu\left(\mathbf{r}\right)^\top d\mathbf{r}d\mathbf{s}
\label{eq:ConstructionIntegral2}
\end{equation}
where $\boldsymbol\Lambda_{x}, \boldsymbol\Lambda_{\theta} \in \mathbb{R}^{n_x \times n_x}$. The spatial IDE model in (\ref{eq:ConvolutionIntegral}) and (\ref{eq:observation equation}) can be represented as linear state-space model of the form  
\begin{equation}
 \mathbf{x}_{t+1}=\mathbf{A}\left(\boldsymbol\theta\right)\mathbf{x}_t+\mathbf{w}_t
\label{eq:StateEquation}
\end{equation}
\begin{equation}
\mathbf{y}_t=\mathbf{C}\mathbf{x}_t+\mathbf{v}_t
\label{eq:ObservationEquation}
\end{equation}
where 
\begin{equation}
 \mathbf{A}\left(\boldsymbol\theta\right)=\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda_{\theta}}
\label{eq:TransitionMatrix}
\end{equation}
and  \\ \\ 
$\mathbf{C}=$
$\begin{pmatrix}
\boldsymbol\phi_{j_0}\left(\mathbf{s}_1\right) & \boldsymbol\phi_{j_0}\left(\mathbf{s}_2\right)  &\cdots &\boldsymbol\phi_{j_0}\left(\mathbf{s}_{n_y}\right) \\
\boldsymbol\psi_{j_0}\left(\mathbf{s}_1\right) & \boldsymbol\psi_{j_0}\left(\mathbf{s}_2\right)  &\cdots &\boldsymbol\psi_{j_0}\left(\mathbf{s}_{n_y}\right) \\
\boldsymbol\psi_{j_0+1}\left(\mathbf{s}_1\right) & \boldsymbol\psi_{j_0+1}\left(\mathbf{s}_2\right)  &\cdots &\boldsymbol\psi_{j_0+1}\left(\mathbf{s}_{n_y}\right) \\
\vdots & \vdots & \ddots& \vdots \\
\boldsymbol\psi_{j}\left(\mathbf{s}_1\right) & \boldsymbol\psi_{j}\left(\mathbf{s}_2\right)  &\cdots &\boldsymbol\psi_{j}\left(\mathbf{s}_{n_y}\right) \\
\end{pmatrix}^{\top}$ \\ \\
\\ \\ and where 
\begin{equation}
 \mathbf{w}_t=\mathbf{\Lambda}_{x}^{-1}\int_{\mathcal{S}}\boldsymbol\mu \left(\mathbf{s}\right)e_t\left(\mathbf{s}\right)d\mathbf{s}
\label{CovofField}
\end{equation}
 is a vector valued, zero-mean normally distributed white noise process with covariance given by (Lemma \ref{lemma:FieldCovariance}, Appendix)
\begin{equation}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\boldsymbol\Pi\mathbf{\Lambda}_{x}^{-\top}
\end{equation}
and
\begin{equation}
 \boldsymbol\Pi=\iint\limits_{\mathcal{S}}\boldsymbol\mu\left(\mathbf{s}\right)\eta\left(\mathbf{s}-\mathbf{r}\right)\boldsymbol\mu\left(\mathbf{r}\right)^{\top}d\mathbf{r}d\mathbf{s}
\label{eq:Pi}
\end{equation}
By substituting the  parameter vector $\boldsymbol\theta$, the  state vector $\mathbf{x_t}$, the vector of kernel basis functions $\boldsymbol\lambda$ and the vector of field basis functions $\boldsymbol\mu$ in (\ref{eq:ConstructionIntegral1}-\ref{eq:Pi}), the state-space representation of the multiresolution approximation of the IDE can be achieved. It should be noted that the complexity of the model is independent of the number of observation locations, providing more flexibility in handling high resolution data. %The model is also capable of being applicable to scenarios with irregular observation locations.

@@

By exploiting properties (\ref{eq:BsplineConvolution}) and (\ref{eq:BsplineInnerProduct}) the  integrals (\ref{eq:ConstructionIntegral1}) and (\ref{eq:ConstructionIntegral2}) can be computed. In order to calculate elements of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Lambda_{\theta}$, scaling and wavelet basis functions should be expanded  in terms of $N_m$ at the appropriate scale. It is important to note that B-spline scaling and wavelet functions possess the following orthogonality properties \cite{Unser1993}: 
 \left\langle \psi_{m;j_1,l_1}(s),\psi_{m;j_2,l_2}(s)\right\rangle =0  \quad \mathrm{for} \quad j_1\neq j_2
\label{PsiPsiOrthogonality}
 \left\langle \phi_{m;j_1,l_1}(s),\psi_{m;j_2,l_2}(s)\right\rangle =0  \quad \mathrm{for} \quad j_1\leq j_2
\label{PhiPsiOrthogonality}
Where the subscript $m$ is introduced to $\phi_{j,l}(s)$ and $\psi_{j,l}(s) $ to indicate the $m$th order B-spline scaling and wavelet functions. In the two dimensional case, the convolution integral can be factorised as follows, e.g. 
\setlength{\arraycolsep}{0.14em}
\begin{IEEEeqnarray}{!l l }
 \phi_{m;j,\mathbf{l}}\left(s_{1},s_{2}\right)\ast \psi^{1}_{m;j,\mathbf{l'}}\left(s_{1},s_{2}\right)&=\left[\phi_{m;j,l_1}\left(s_{1}\right)\ast\phi_{m;j,l'_1}\left(s_{1}\right)\right] \nonumber \\
&\times \left[ \phi_{m;j,l_2}\left(s_{2}\right) \ast \psi_{m;j,l'_2}\left(s_{2}\right)\right] \nonumber \\
&
The preceding factorisation can be also applied to the 2D-inner product integral. Non-zero quantities can  be then expressed by
\setlength{\arraycolsep}{0.14em}
\begin{IEEEeqnarray}{!l l }
\left\langle \phi_{m;j,\mathbf{l}}\left(s_{1},s_{2}\right),\phi_{m;j,\mathbf{l'}}\left(s_{1},s_{2}\right)\right\rangle &=\left\langle \phi_{m;j,l_1}\left(s_{1}\right),\phi_{m;j,l'_1}\left(s_{1}\right)\right\rangle \nonumber \\
& \times\left\langle \phi_{m;j,l_2}\left(s_{2}\right) , \phi_{m;j,l'_2}\left(s_{2}\right)\right\rangle \nonumber \\
&
likewise we can compute 
$\left\langle  \psi^{1}_{m;j,\mathbf{l}}\left(s_{1},s_{2}\right),\psi^{1}_{m;j,\mathbf{l'}}\left(s_{1},s_{2}\right)\right\rangle$
, \\
$\left\langle  \psi^{2}_{m;j,\mathbf{l}}\left(s_{1},s_{2}\right),\psi^{2}_{m;j,\mathbf{l'}}\left(s_{1},s_{2}\right)\right\rangle$ and \\
$
\left\langle \psi^{3}_{m;j,\mathbf{l}}\left(s_{1},s_{2}\right), \psi^{3}_{m;j,\mathbf{l'}}\left(s_{1},s_{2}\right)\right\rangle $. \\ \\
Using these expansions it becomes straightforward to calculate the integrals (\ref{eq:ConstructionIntegral1}) and (\ref{eq:ConstructionIntegral2}). In this work, 4th order cardinal B-spline scaling and wavelet functions are used therefore, to compute the integrals (\ref{eq:ConstructionIntegral1}) and (\ref{eq:ConstructionIntegral2})  the $8$th and $12$th order B-spline values at integer points are  required and can be found in \cite{Goswami1999}.
\section{Estimation of the Multiresolution Spatio-temporal IDE Model}
A state-space representation of the IDE model allows the use of the Expectation Maximisation (EM) algorithm to infer both the kernel and the spatial field from the noise corrupted spatio-temporal data. In this context, the EM algorithm is a two part iterative algorithm that yields the maximum likelihood estimate of the spatial mixing kernel parameters and the posterior distribution of the field weights. In the M-step a lower bound $\mathcal{Q}\left(\boldsymbol\theta,\boldsymbol\theta'\right)$ of the complete-data log-likelihood  $p\left(\mathbf{X},\mathbf{Y} \mid \boldsymbol\theta \right)$, is maximised under the assumption of the system parameters being $\boldsymbol\theta'$ \cite{McLachlan1997}. In the E-step, given an estimate of $\boldsymbol\theta$, filtering methods can be used to determine the posterior state sequence. Here we compare three techniques namely based on  the Rauch Tung Striebel (RTS) smoother \cite{Gibsona2005}, Ensemble Kalman Smoother (EnKS) \cite{Evensen2003} and a heuristic separable RTS to solve the joint estimation problem. For the first method we modify the framework proposed in \cite{Dewar2009} by computing the required elements once instead of at each iteration, the modified algorithm is computationally less demanding. see Appendix ...

@@

In this section, we compare the performance of the different algorithms mentioned above on a synthetic example .....This section demonstrates the identification of the proposed IDE model. Primarily, the MRA of a synthetic model is demonstrated, which is able to represent the system at different spatial scales. In addition, the complexity of the field decomposition is shown to be independent to the number and spatial regularity of the observation locations. This allows the construction of parsimonious models using spatially dense observations, and demonstrates the utility of the approach to situations where the observations are not arranged on a regular lattice, i.e. sensor networks \cite{Sahu2008}.

@@

\section*{Reducing the computational complexity}
The state-space representation of the spatio-temporal IDE \eqref{eq:StateEquation} and \eqref{eq:ObservationEquation} has a quadratic $\mathcal Q$-function in a form of \cite{Dewar2009}
\mathcal{Q}\left(\boldsymbol\theta,\boldsymbol\theta'\right)=2 \boldsymbol\theta^\top\boldsymbol\upsilon -\boldsymbol\theta^\top\boldsymbol\Upsilon\boldsymbol\theta
\label{eq:QFunction}
which has a maximum at
 \boldsymbol\theta=\boldsymbol\Upsilon^{-1}\boldsymbol\upsilon
where
 \boldsymbol\Upsilon=\sum_{i,j,k,m=1}^{n_x}\left[ \Xi_1\right] _{ij}\boldsymbol\gamma^{\left(k,j\right)}\left[\boldsymbol\Lambda_{x}^{-1}\Sigma_{w}^{-1} \boldsymbol\Lambda_{x}^{-1}\right]_{km} \boldsymbol\gamma^{\left(m,i\right)\mathsf{T}}
 \boldsymbol \upsilon=\sum_{i,j=1}^{n_x}\left[\Xi_0\Sigma_{w}^{-1}\boldsymbol\Lambda_{x}^{-1} \right]_{ij}\boldsymbol\gamma^{\left(j,i\right)} 
where
 \boldsymbol\gamma^{(i,j)}=\int_{\mathcal S}\left[ \mu(s)\right]_i\left[ \mathbf M(s)\right] _{:j}ds
and $\left[ \mathbf M(s)\right] _{:j}$ denotes the $j$th column of $\mathbf M(s)$. The definitions of the $\Xi_0$ and $\Xi_1$ are the same as in \cite{Dewar2009}. Note that constant values with respect to $\boldsymbol \theta$ are ignored in \eqref{eq:QFunction}.

@@

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
  \newpage
% \bibliographystyle{unsrt} 
% \bibliography{MRAIDE}

\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
 \bibliography{IEEEabrv,MRAIDE}
